A data-driven, pixel-level model to improve the precision of \Kepler\ photometry

### 2013-10-18 documents/plm.tex
objective function, optimization, notes to self

As mentioned above, in the context of this project,
  a model is something that can predict data given settings of some parameters,
  and an objective function that can be used to find best values for those parameters.
We prefer to work in a probabilistic mode in which the objective function can be justified
  in terms of a likelihood (probability for the data given the model)
  or a posterior probability density function (likelihood times a prior pdf).
The PLM treats the noise in a \Kepler\ pixel $m$ at time $n$ as Gaussian (normally distributed)
  with mean zero and known variance $\sigma^2_{mn}$,
  and the expectation value for each \Kepler\ data point $I_{mn}$ as being
  predictable from a linear combination of data points $I_{m'n}$,
  where $m'$ is from the set of non-overlapping pixels $m'\in\set{M}_m$.
This linear model has parameters (linear coefficients) $a_{mm'}$.
\todo{foo:}
\begin{eqnarray}
I_{mn}         &=& I^{\ast}_{mn} + e_{mn}
\\
I^{\ast}_{mn}  &=& \sum_{m'\in\set{M}_m} a_{mm'}\,I_{m'n} + \sum_{k=1}^{K} b_{mk}\,d_{kn}
\\
p(e_{mn})      &=& N(e_{mn}\given 0,\sigma^2_{mn})
\quad,
\end{eqnarray}
where HOGGGGGGGGGGGGGGGGGGGGGGGG
\begin{eqnarray}
\chi^2_{mn}    &=& \sum_{n'\in\set{N}_n} \frac{[I_{mn'} - I^{\ast}_{mn'}]^2}{\sigma^2_{mn'}}
                 + \sum_{m'\in\set{M}_m} \frac{a_{mm'}^2}{\Sigma^2}
\\
I^{\ast}_{mn'} &=& \sum_{m'\in\set{M}_m} a_{mm'}\,I_{m'n'}
\quad,
\end{eqnarray}
where \todo{bar}

We train the model how?

We choose the strength of the regularization how?

We do this for every pixel and every time step.
Implementation issues.

@@

Lots of quantitative and qualitative choices that ought to be explored.
These include what and what and what and what?

>>>>

As mentioned above, in the context of this project,
  a model is something that can predict data given settings of some parameters,
  and an objective function that can be used to find best values for those parameters.
We prefer to work in a probabilistic mode in which the objective function can be justified
  in terms of a likelihood (probability for the data given the model)
  or a posterior probability density function (likelihood times a prior pdf).
The PLM treats each \Kepler\ data point $I_{mn}$ as being
  predictable from a linear combination of data points $I_{m'n}$,
  where $m'$ is from the set of non-overlapping pixels $m'\in\set{M}_m$.
\begin{eqnarray}
I_{mn}         &=& I^{\ast}_{mn} + e_{mn}
\\
I^{\ast}_{mn}  &=& \sum_{m'\in\set{M}_m} a_{mm'}\,I_{m'n} + \sum_{k=1}^{K} b_{mk}\,d_{kn}
\quad,
\end{eqnarray}
where $I^{\ast}_{mn}$ is the prediction (by the model) for data point $I_{mn}$,
  $e_{mn}$ is a noise contribution or residual away from the prediction,
  and the $a_{mm'}$ and $b_{mk}$ are the parameters (linear coefficients of the prediction).
If we presume that the residuals away from the prediction are normally distributed with zero mean
  and known variance,
  likelihood optimization reduces to chi-squared minimization.
We add to the standard chi-squared definition a regularization term
  (equivalent to multiplying the likelihood by a Gaussian prior pdf)
  that penalizes large absolute values for the the coefficients $a_{mm'}$:
\begin{eqnarray}
\chi^2_{mn}    &=& \sum_{n'\in\set{N}_n} \frac{[I_{mn'} - I^{\ast}_{mn'}]^2}{\sigma^2_{mn'}}
                 + \sum_{m'\in\set{M}_m} \frac{a_{mm'}^2}{\Sigma^2}
\\
I^{\ast}_{mn'} &=& \sum_{m'\in\set{M}_m} a_{mm'}\,I_{m'n'} + \sum_{k=1}^{K} b_{mk}\,d_{kn'}
\quad,
\end{eqnarray}
where the $\sigma^2_{mn'}$ are the (presumed known) individual-pixel noise variances,
  and $\Sigma^2$ sets the strength of the regularization (or width of the prior pdf).
\todo{DFM: We set or ``know'' the variances how?
  I think we should use the inverse of the median of the reported inverse variances.
  Also, we set the prior strength to what value?
  Heuristic justification of the latter.}

The train-and-test idea comes into the objective function $\chi^2_{mn}$:
When we are computing the objective function $\chi^2_{mn}$,
  we are using only time points $n'\in\set{N}_n$ that don't overlap the target (or test) time point $n$.
We train the model---that is, set the parameters $a_{mm'}$ and $b_{mk}$---%
  by choosing the full set of parameters that jointly minimize the objective function $\chi^2_{mn}$.
Importantly, and perhaps surprisingly, this is a \emph{different objective function}
  for every target data point (pixel value to be predicted) $I_{mn}$,
  and we have to do a \emph{independent optimization} of the parameters for every pixel value we want to predict.
Since there are about $10^{11}$ total pixel values in the \Kepler\ \todo{what imaging?} data archive,
  this represents one fuck of a lot of optimizations.

For this reason---\emph{so many optimizations}---%
  we do \todo{some crazy clever stuff} to make this work fast;
  \todo{DFM implementation notes here}.

@@

Lots of quantitative and qualitative choices that ought to be explored.
These include what and what and what and what?
How \emph{ought we} make all these choices?
